Using Hybrid Machine Learning for Fundamental Stellar Parameter Inference

### 2024-05-29 paper/paper2/paper.tex
Gully edits in Overleaf

\begin{abstract}
    The advent of machine learning methods has revolutionized 
    numerous scientific methods, among which is spectroscopic 
    inference. Studies can and have shown that machine learning 
    methods prove effective in inferring what we call
    fundamental stellar parameters. These include $T_{\mathrm{eff}}$,
    $\log(g)$, and [Fe/H] for the purposes of this study but can 
    also encompass other parameters such as [$\alpha$/Fe], C/O, and
    $f_{sed}$. While current methods tend to treat spectra purely
    from a data perspective, we propose to use a hybrid machine 
    learning (HML) method that uses the semi-empirical approach 
    of treating spectra as a set of spectral lines put forth by 
    \texttt{blase} \citep{blase}, combined with traditional linear interpolation 
    as well as Bayesian optimization. We interpolate the four 
    line properties ($\mu'$, $\ln(a)$, $\sigma$, $\gamma$) of 
    128,723 unique spectral lines across a subset of the PHOENIX
    synthetic model grid \citep{PHOENIX} to create the PHOENIX generator, allowing
    us to reconstruct spectra at any continuously valued point 
    lying within the domain of the PHOENIX subset, and even reconstruct 
    batches of spectra in pseudo-parallel fashion, with the speedup 
    exceeding 20 times once more than 60 spectra are reconstructed. 
    We then combine the PHOENIX generator with Bayesian optimization to 
    efficiently infer fundamental stellar parameters. This study is 
    however a proof of concept, as we have not yet included various
    features that would be necessary for a full-fledged tool for 
    spectroscopic inference, however we show that semi-empirical 
    HML methods are a viable, albeit fledgling alternative to 
    more traditional approaches.
\end{abstract}

@@

The challenge of extracting as much information as possible from
a stellar spectrum to get back the properties that lie at the root of it all. 
However, this has proven to be no easy task. Spectroscopic surveys such as 
APOGEE develop their own in-house pipelines to extract stellar parameters
from spectra (ASPCAP in the case of APOGEE), and these pipelines are fairly
effective \citep{APOGEE, ASPCAP}. However, they all work on the core assumption that the data to be 
analyzed is a list of pixels. In addition, the pipelines that surveys 
develop are typically closed-source and usually limited to the scope of the 
survey itself, causing a sort of fragmentation. This motivates the 
development of a more universal, open-source tool that can do what spectral 
analysis pipelines do, but in an alternative way.

Multiple efforts have been made in this direction, treating spectra in 
different ways. The standard practice is to treat the wavelength and flux as 
simply two arrays and implement any from a myriad of different algorithms to 
inference fundamental stellar parameters. Others take a more mathematical 
approach by decomposing spectra into some eigenbasis and treating it as a 
regression problem, such as \texttt{starfish} \citep{starfish}. However, \texttt{blase} takes 
a different approach. It treats spectra not as a set of pixels or a set of 
eigenbasis coefficients but as a set of spectral lines, specifically Voigt 
profiles. There is no metric of this approach being better or worse than the 
others mentioned so far, but it is a different paradigm and is worth 
exploring as a semi-empirical tool. This study aims to take advantage of 
\texttt{blase}'s ability to change a spectrum's basis from flux versus 
wavelength into spectral line versus spectral line properties, and use that 
to develop a robust tool for stellar parameter inference.

However, this is only a proof of concept. Here we develop a tool that 
represents a very basic realization of the idea of line-by-line inference. 
We make no claims as to the performance of this tool, but rather aim to show 
that the idea is sound and able to be implemented, even if rudimentary, and 
that with further research and development, this tool has the potential to 
become useful for spectral inference.

We chose the PHOENIX synthetic spectral model grid as the basis for this 
study, because it is fairly well-known. While we plan to extend this tool to 
leverage information across multiple model grids in the future, both for 
accuracy and to increase the scope to include substellar objects with grids
such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, for now we limit 
our scope to a subset of the PHOENIX grid.

Our study is divided into three main parts. First, we preprocess the PHOENIX
subset and detect spectral lines with \texttt{blase}. Second, we interpolate
the spectral line parameters, enabling us to evaluate the PHOENIX grid at 
any point within the bounds of the subset. Lastly, we use Bayesian optimization 
to power our inference algorithm and test its performance. An overview of this
process is presented in Figure 1.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure1}
    \caption{Overview of the process used in this study.}
\end{figure*}

\section{Cloning the PHOENIX Model Grid}
\subsection{The PHOENIX Subset}
For the purposes of this study, we did not consider the full range of the
PHOENIX synthetic spectral model grid. Instead, we focused on a subset of 
the grid, whose parameter ranges are given in Table 1.

@@

\subsection{Preprocessing with \texttt{gollum}}
First, the PHOENIX subset was accessed directly from the PHOENIX website
using \texttt{gollum}'s download option \citep{gollum}. The spectra were then put through a 
three-step preprocessing pipeline.
\begin{enumerate}
    \item \textit{Blackbody Division}: Since the $T_{\mathrm{eff}}$ of each spectrum 
    is known, the according blackbody spectrum was divided out.
    \item \textit{Percentile Normalization}: The spectra were normalized by dividing
    them by their 99th percentile.
    \item \textit{Continuum Normalization}: The spectra were further normalized by 
    dividing them by a \nth{5} order polynomial continuum.
\end{enumerate}
Mathematically, we can express the preprocessing as follows:
\begin{gather}
    \mathsf{\bar{S}} = \dfrac{\mathsf{S}}{\mathsf{B}\mathsf{Q}_5P_{99}}
\end{gather}
where $\mathsf{\bar{S}}$ is the preprocessed spectrum, $\mathsf{S}$ is the 
original spectrum,  $\mathsf{B}$ is the blackbody spectrum, $\mathsf{Q}_n$ 
is the $n^\mathrm{th}$ order polynomial continuum fit, and $P_n$ is the 
$n^\mathrm{th}$ percentile function. Arithmetic operations between arrays 
are assumed to be elementwise in all notation from here on out.

\subsection{Line Identification with \texttt{blase}}
The next step was to convert the PHOENIX subset into an interpretable format.
We wanted to represent the spectra as a list of spectral lines rather than a
list of fluxes. This was done using \texttt{blase}, which detects spectral 
lines as Voigt profiles and tunes the profiles to mimic the original 
PHOENIX spectrum with back propagation. Four parameters were optimized: the 
line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the 
Lorentzian width $\gamma$. The optimization used the Adam optimizer with a
learning rate of 0.05 over 100 epochs \citep{adam}. In addition, we limited two custom 
parameters: wing cut to 6000 and prominence to 0.005. Wing cut is a parameter
that determines the extent of the Voigt profile to evaluate, saving
computational resources by not evaluating small numbers. Prominence sets a
lower limit for the amplitude of detected lines, which also saves resources 
by disregarding shallow lines. In short, larger values for wing cut and 
smaller values for prominence both increase the accuracy of using
\texttt{blase}'s line detection and optimization at the expense of an
increased computational cost. In addition, it should be mentioned that
\texttt{blase} uses the pseudo-Voigt approximation, which saves on 
computational cost while remaining accurate to about 1\% \citep{pseudovoigt1, pseudovoigt2}. It is a 
weighted average of a Gaussian and Lorentzian as opposed to a convolution.
\texttt{blase}'s pseudo-Voight profile implementation uses the following:
\begin{subequations}
\begin{gather}
    \tilde{\mathsf{V}}_\mu(\lambda) = a\left[\eta \mathbf{L}(\lambda - \mu'; f) + (1 - \eta)\mathbf{G}(\lambda - \mu'; f)\right]\\
    \eta = \sum_{n=1}^3 \mathbf{u}_n\left(\dfrac{2\gamma}{f}\right)^n\\
    f = 32\sum_{n=0}^5 \mathbf{v}_n \left(\sqrt{2\ln(2)}\sigma\right)^{5-n}(\gamma)^n\\
    \nonumber \mathbf{u} = \mqty[1.36603 \\ -0.47719 \\ 0.11116] \quad \mathbf{v} = \mqty[1 \\ 2.69269 \\ 2.42843 \\ 4.47163 \\ 0.07842 \\ 1]
\end{gather}
\end{subequations}
where $\mathbf{L}$ and $\mathbf{G}$ are abbreviations for Lorentzian and 
Gaussian profiles, respectively. Notice that we use $\mu'$ instead of $\mu$ 
in the formula. This is because \texttt{blase} allows the line center to shift 
slightly during optimization, and it is this shifted center which is used in 
computation. The individual Voigt profiles are still indexed by $\mu$ for 
cross-model line identification, explained in the next section. Once optimization 
was complete, the list of identified lines, present in the state dictionary of 
\texttt{blase}'s model, was saved to a \texttt{.pt} file for each PHOENIX subset 
grid point \citep{pytorch}. The total disk space these files took up is 465 MB.

\section{Interpolating Manifolds}
\subsection{Cross-Model Line Identification}
Earlier we mentioned that \texttt{blase} tuned the line centers of
detected lines. This means that from one PHOENIX spectrum to the next, the
same line could have a slightly different line center. Since the goal of this
study is to interpolate the properties of each line, we needed to identify the
presence of a particular line across the PHOENIX subset. We decided to do
this by using the line centers $\mu$ of the detected lines pre-optimization. 
Now with each spectral line indexed by $\mu$, we had four parameters to interpolate:
$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. Note that since we are dealing with 
the parameters of Voigt profiles, we can see in Equation 2 that even if the 
interpolation method is linear, the actual result of interpolation varies 
non-linearly.

There is also a second issue, and that is that spectral lines were often
only detected in some spectra from the PHOENIX subset. In Figure 2, we show
that different grid points sport differing counts of detected spectral lines.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure2}
    \caption{Number of detected spectral lines at each grid point of 
    a slice of the PHOENIX subset at solar metallicity. We can see that
    the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
    Also note that from $T_{\mathrm{eff}} = 7000$ K onward, the spacing 
    between grid points increases from 100 K to 200 K.}
\end{figure*}
This means that in theory, interpolation would be inaccurate in regions where 
a line does not appear, and also breaks in practice because \texttt{scipy}'s 
\texttt{RegularGridInterpolator} relies on a rectilinear grid, which we would not 
have if regions of grid points were missing.

To solve this, we artificially
populated missing grid points with log-amplitudes of -1000, which retained
interpolator stability by not being an infinity, but also essentially 
nullified the line when evaluated. An example of the appearance of missing 
sections in heatmaps where a line does not appear is shown in Figure 3. 
In total, across the entire PHOENIX subset, \texttt{blase} detected 
128,723 individual spectral lines, meaning we have that many manifolds to 
interpolate.

@@

\subsection{Continuously Evaluable Manifolds}
For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$,
$\log(g)$, and [Fe/H], and the output was a list of four parameters, 
$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. For each line, one of these 
interpolator objects was created using linear interpolation, and these 
interpolators were aggregated into a single list, which was when pickled 
and written to a \texttt{.pkl} file. These interpolators generate multiple 
manifolds with the following mapping:
\begin{gather}
    \mqty[T_{\mathrm{eff}} \\ \log(g) \\ \mathrm{[Fe/H]}] \rightarrow \mqty[\mu' \\ \ln(a) \\ \sigma \\ \gamma]
\end{gather}
These interpolators could now be evaluated at any point lying within the 
domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset 
into a continuous one. With the given size of the PHOENIX subset, the 
interpolator list takes up 13.2 GB of disk space. This evaluation 
is able to reconstruct an existing PHOENIX spectrum or alternatively 
interpolate a new spectrum within the domain of the PHOENIX subset,
we call this the PHOENIX generator. In Figure 4, we show the same 
spectral line as in Figure 3, but now supersampled using the PHOENIX
generator evaluated over the same slice.

The spectral reconstruction process is done by iterating over the 
PHOENIX generator, evaluating each interpolator at the given coordinates, 
then reshaping the data into the same format that PyTorch uses for 
state dictionaries. During the iteration, if the interpolated
log-amplitude of the line is less than -100, the line is excluded from the 
state dictionary. Why -100 instead of -1000? This is because near the 
areas that we artificially populated, the manifold experiences artifacts due 
the sudden decline. We have to anticipate that decline and cut off the 
behavior in advance. This saves on computational time. Then, the state
dictionary is fed into \texttt{blase}'s \texttt{SparseLinearEmulator},
which reconstructs the spectrum by constructing a forward model based on 
the input state dictionary. Any \texttt{nan} values are set to 1 (which we 
can do because the spectra are all normalized), and the spectrum is 
complete.

@@

\section{Bayesian Inference and Testing}
\subsection{Spectral Reconstruction Time}
The typical use case for the PHOENIX generator may be to batch 
reconstruct spectra from an array of input coordinates. Therefore, there
is some motivation to attempt to reduce the computational cost of this
procedure to tractable levels. We evaluate the computational time needed
to use the PHOENIX generator in two distinct ways. First, for a single
input, which would be relevant in serial applications. Second, for an array
of multiple inputs, which would be relevant in parallel applications.
The \texttt{RegularGridInterpolator} API allows for the passing in of 
an entire array of input coordinates to be evaluated at once, however 
using \texttt{blase} to reconstruct the spectrum from our interpolated
state dictionary is done serially, leading to what is actually more of 
a psuedo-parallel evaluation. Performance results are shown in Figure 5,
and we can see that the pseudo-parallel implementation is much faster.

@@

\subsection{Inference Algorithm}
We elected to use Bayesian optimization as the inference algorithm, specifically 
the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library \citep{skopt}. 
This algorithm uses a Gaussian Process to model the objective function, which in 
this case was the RMS (Root-Mean-Square) loss between the interpolated spectrum 
$\mathsf{M}$ and the true spectrum $\mathsf{D}$, defined as:
\begin{gather}
    \mathcal{L} = \ev{(\mathsf{M} - \mathsf{D})^{\circ 2}}^{1/2}
\end{gather}
The optimizer was configured to first run 100 random 
evaluations to seed the surrogate model, then run 20 more evaluations now guided 
by the surrogate model. This totals to 120 evaluations, which was deemed 
sufficient for this study. One inference run takes us on average, just under 
7.5 minutes to complete.

\subsection{Bayesian Optimizer Performance}
To test the performance of the inference algorithm, we used the PHOENIX subset
itself. At first glance, this may seem like a useless endeavor, as the 
PHOENIX generator has indeed memorized the PHOENIX subset, being able to
reconstruct a PHOENIX spectrum when evaluated at that grid point. However,
that is precisely what allows us to use the PHOENIX subset as test data.
The Bayesian optimizer's surrogate model is seeded by random continuously 
sampled generator evaluations within the search space, \textit{not} grid 
points of the PHOENIX subset, meaning the surrogate model has no `memorization' 
to speak of. If the optimizer had been an iterative grid refinement search, 
this would not have been possible, because grid methods rely on just that, 
the grid.

We know that $T_{\mathrm{eff}}$ tends to be fairly well-constrained prior to 
even looking at the spectra, so when testing the inference algorithm, we limited
its search space to only include $T_{\mathrm{eff}}$ which lay within 500 K of the
true value on either side. $\log(g)$ and [Fe/H] were allowed to vary freely. 
The test sample consisted of 9 unique $T_{\mathrm{eff}}$ values, 3 unique $\log(g)$ 
values, and 2 unique [Fe/H] values, totaling 54 unique spectra. $T_{\mathrm{eff}}$
ranged from 3000 K to 11000 K in increments of 1000 K, $\log(g)$ ranged from 2 to 6
in increments of 2, and [Fe/H] ranged from -0.5 to 0 in increments of 0.5.

@@

\section{Final Considerations}
The computer used for this study, Triton, has the following specifications:
\begin{table}[h!]
    \centering
    \begin{tabular}{ll}
        \hline
        CPU & AMD EPYC 7513\\
        RAM & 256 GB\\
        GPU & Nvidia A100 40GB ($\times$2)\\
        \hline
    \end{tabular}
    \caption{Triton is owned by Caroline Morley and is available to
    members of her research group. It was used for all computations, but not 
    for generating visualizations. The EPYC 7513 is a 32c/64t CPU with a 
    boost clock of 3.65 GHz. The A100 has 6912 CUDA cores.}
\end{table}

It should be noted that this study represents a proof of concept, and that 
there are numerous design considerations that could be improved upon with 

@@

    study was just that, a subset. The full PHOENIX grid not only expands 

@@

    but also includes the alpha element abundance parameter, which we 

>>>>

\begin{abstract}
    The advent of machine learning methods is revolutionizing 
    numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference. Machine learning applied to spectroscopy may be expected to infer fundamental stellar parameters at greater speed, accuracy, and/or precision than conventional techniques. Here we introduce a new scalable ML technique to infer $T_{\mathrm{eff}}$,
    $\log(g)$, and [Fe/H] parameters, and the extensible method holds promise for other properties such as [$\alpha$/Fe], C/O, and
    $f_{sed}$. Our method contrasts with traditional data-driven
    approaches by introducing hybrid machine 
    learning (HML), a semi-empirical approach 
    to treating spectra as differentiable spectral lines described in 
    \texttt{blase} \citep{blase}.  
    \authorcomment1{Say how many spectra you analyze in the ``PHOENIX subset''.  State that an outcome of this work is the precomputed training.}
    We combine traditional linear interpolation 
    and Bayesian optimization to create a manifold mapping between 
    line properties ($\mu'$, $\ln(a)$, $\sigma$, $\gamma$) to fundamental paramerts.  of 
    128,723 unique spectral lines across a subset of the PHOENIX
    synthetic model grid \citep{PHOENIX} to create the PHOENIX generator, allowing
    us to reconstruct spectra at any continuously valued point 
    lying within the domain of the PHOENIX subset, and even reconstruct 
    batches of spectra in pseudo-parallel fashion, with the speedup 
    exceeding 20 times once more than 60 spectra are reconstructed. 
    We then combine the PHOENIX generator with Bayesian optimization to 
    efficiently infer fundamental stellar parameters. This study is 
    however a proof of concept, as we have not yet included various
    features that would be necessary for a full-fledged tool for 
    spectroscopic inference, however we show that semi-empirical 
    HML methods are a viable, albeit fledgling alternative to 
    more traditional approaches.
\end{abstract}

@@

The challenge lies in extracting as much information as possible from
a stellar spectrum to get back the stellar fundamental properties. 
Spectroscopic surveys such as APOGEE develop their own in-house pipelines to extract stellar parameters
from spectra (ASPCAP in the case of APOGEE), and these pipelines appear
effective \citep{APOGEE, ASPCAP}. However, they all work on the core assumption that the data to be 
analyzed is a list of pixels. Analysis pipelines may be closed-source or limited to the scope of the survey itself, causing a sort of siloing effect among surveys. This motivates the 
development of a more universal, instrument-agnostic open-source framework that can apply broadly to a range of spectral observations with relatively little tuning.

Multiple efforts have been made in this direction, treating spectra in 
different ways. The standard practice is to treat the wavelength and flux as 
simply two data arrays and bespoke algorithms tailored to a small number of well-calibrated spectral lines to obtain fundamental stellar parameters and chemical compositions. Other whole-spectrum fitting abstractions decompose model spectra into an eigenbasis, simplifying the data-model comparision stage nto a tractable regression, such as \texttt{starfish} \citep{starfish}. However, \texttt{blase} takes 
an intermediate approach \authorcomment1{cite}. 
It treats spectra not as a set of pixels or a set of  eigenbasis coefficients but as a set of spectral lines, specifically Voigt profiles. Each of these approaches has tradeoffs, but one key advantage places blase in a new category: the ability to fit the non-linear spectral line parameters with automatic differentiation.  This autodiff paradigm positions blase a promising semi-empirical tool.\authorcomment1{consider citing exojax and wobble, which also use autodiff for spectra}
\authorcomment1{Say a bit more on what blase paper 1 limitations-- only demonstrated on a few spectra with hand-chosen fiducial models, not scaled out to the entire grid.}

Ideally this process could be end-to-end differentiable, with all spectral inference code living inside a single autodiffable machine learning framework, like \texttt{PyTorch} or \texttt{JAX}. Here we separate the problem into two pieces.  First, the scaling out of the blase method to over 1000 precomputed synthetic spectral model clones, yielding a downloadable cache of pretrained machine learning models with over xx0,000 spectral lines.  This product alone serves as a useful asset to the community, since it can serve as a lookup for individual lines. Second, we explore the properties of the manifold mapping between fundamental stellar properties and the underlying spectral lines.  Finally, we show how reconstructions of the spectra through the manifold can be adapted for end-to-end spectral inference.
Extensions to this underlying framework have the potential to be the best in class with further research and development.

We chose the widely-adopted PHOENIX synthetic spectral model grid \authorcomment{cite?} as the basis for this study. The framework can be straightforwardly extended to any other model grid in the future, including substellar atmosphere grids such as Sonora \citep{bobcat, cholla, diamondback, elfowl}, for now we limit our scope to a subset of the PHOENIX grid.

\authorcomment1{whoops, I repeated this above.  Key idea is to put some numbers here.}
Our study is divided into three main parts. First, we preprocess the PHOENIX
subset and detect spectral lines with \texttt{blase}. Second, we interpolate
the spectral line parameters, enabling us to evaluate the PHOENIX grid at 
any point within the bounds of the subset. Lastly, we use Bayesian optimization to power our inference algorithm and test its performance. An overview of this process is presented in Figure 1.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure1}
    \caption{Overview of the process used in this study.}
\end{figure*}

\section{Cloning the PHOENIX Model Grid}
\subsection{The PHOENIX Subset}
For the purposes of this study, we did not consider the full range of the
PHOENIX synthetic spectral model grid. Instead, we focused on a subset of 
the grid, with the parameter ranges given in Table 1.
\authorcomment1{state the number of spectra in the subset if you have not already done so.  optionally state how many teff, logg, fe/h spacings.}

@@

\subsection{Preprocessing with \texttt{gollum}}
First, the PHOENIX subset was programmatically retrieved with \texttt{gollum} \citep{gollum}. The spectra were then put through a 
three-step preprocessing pipeline, identical to \authorcomment1{blase paper 1}.
\begin{enumerate}
    \item \textit{Blackbody Division}: Since the $T_{\mathrm{eff}}$ of each spectrum 
    is known, the according blackbody spectrum was divided out.
    \item \textit{Percentile Normalization}: The spectra were normalized by dividing
    them by their 99th percentile.
    \item \textit{Continuum Normalization}: The spectra were further normalized by 
    dividing them by a \nth{5} order polynomial continuum.
\end{enumerate}
Mathematically, we can express the preprocessing as follows:
\begin{gather}
    \mathsf{\bar{S}} = \dfrac{\mathsf{S}}{\mathsf{B}\mathsf{Q}_5P_{99}}
\end{gather}
where $\mathsf{\bar{S}}$ is the preprocessed spectrum, $\mathsf{S}$ is the 
original spectrum,  $\mathsf{B}$ is the blackbody spectrum, $\mathsf{Q}_n$ 
is the $n^\mathrm{th}$ order polynomial continuum fit, and $P_n$ is the 
$n^\mathrm{th}$ percentile function. Arithmetic operations between arrays 
are assumed to be elementwise in all notation from here on out.

\subsection{Line Identification with \texttt{blase}}
The next step was to convert the PHOENIX subset into a physically interpretable intermediate representation: a table of spectral lines properties rather than a list of fluxes. This was done using \texttt{blase}, which detects spectral lines as Voigt profiles and tunes the profiles to mimic the original PHOENIX spectrum with back propagation. 
Four parameters were optimized: the 
line center $\mu$, the log-amplitude $\ln(a)$, the Gaussian width $\sigma$, and the 
Lorentzian width $\gamma$. The optimization used the Adam optimizer with a
learning rate of 0.05 over 100 epochs \citep{adam}. In addition, we limited two custom 
parameters: wing cut to 6000 and prominence to 0.005. Wing cut is a parameter
that determines the extent of the Voigt profile to evaluate, saving
computational resources by not evaluating negligible line wings. Prominence sets a
lower limit for the amplitude of detected lines, which also saves resources 
by disregarding shallow lines. In short, larger values for wing cut and 
smaller values for prominence both increase the accuracy of using
\texttt{blase}'s spectral reconstruction at the expense of an
increased computational cost. In addition, it should be mentioned that
\texttt{blase} uses the pseudo-Voigt approximation, which saves on 
computational cost while remaining accurate to about 1\% \citep{pseudovoigt1, pseudovoigt2}. It is a 
weighted average of a Gaussian and Lorentzian as opposed to a convolution.
\texttt{blase}'s pseudo-Voight profile implementation uses the following:
\begin{subequations}
\begin{gather}
    \tilde{\mathsf{V}}_\mu(\lambda) = a\left[\eta \mathbf{L}(\lambda - \mu'; f) + (1 - \eta)\mathbf{G}(\lambda - \mu'; f)\right]\\
    \eta = \sum_{n=1}^3 \mathbf{u}_n\left(\dfrac{2\gamma}{f}\right)^n\\
    f = 32\sum_{n=0}^5 \mathbf{v}_n \left(\sqrt{2\ln(2)}\sigma\right)^{5-n}(\gamma)^n\\
    \nonumber \mathbf{u} = \mqty[1.36603 \\ -0.47719 \\ 0.11116] \quad \mathbf{v} = \mqty[1 \\ 2.69269 \\ 2.42843 \\ 4.47163 \\ 0.07842 \\ 1]
\end{gather}
\end{subequations}
where $\mathbf{L}$ and $\mathbf{G}$ are abbreviations for Lorentzian and 
Gaussian profiles, respectively. Notice that we use $\mu'$ instead of $\mu$ 
in the formula. This is because \texttt{blase} optionally allows the line center to shift 
slightly during optimization, and it is this shifted center which is used in 
computation. The individual Voigt profiles are still indexed by $\mu$ for 
cross-model line identification, explained in the next section. Once optimization 
was complete, the list of identified lines, was saved to a so-called ``state dictionary'' a common representation for pre-trained machine learning models that can be stored and reopened from this static cache.  These are stored as the \texttt{.pt} file format for each of the $N=xx,000$ PHOENIX subset 
grid point \citep{pytorch}. The total disk space these files took up is 465 MB.\authorcomment1{Restate the number of grid points for good measure, }.

\section{Interpolating Manifolds}
\subsection{Cross-Model Line Identification}
As previously mentioned, \texttt{blase} tunes the line centers of
detected lines. This means that from one PHOENIX spectrum to the next, the
same line could have a slightly different line center. Since the goal of this
study is to interpolate the properties of each line, we needed to identify the
presence of a particular line across the PHOENIX subset, associating the same line with every occurrence. We decided to do
this by using the line centers $\mu$ of the detected lines pre-optimization. 
Now with each spectral line indexed by $\mu$, we had four parameters to interpolate:
$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. Note that since we are dealing with 
the parameters of Voigt profiles, we can see in Equation 2 that even if the 
interpolation method is linear, a final spectral reconstruction will vary 
non-linearly in the flux dimension.

There is also a second issue.  Spectral lines were often
only detected in some spectra from the PHOENIX subset. In Figure 2, we show
that different grid points sport differing counts of detected spectral lines.
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figure2}
    \caption{Number of detected spectral lines at each grid point of 
    a slice of the PHOENIX subset at solar metallicity. We can see that
    the number of detected lines decreases with increasing $T_{\mathrm{eff}}$.
    Also note that from $T_{\mathrm{eff}} = 7000$ K onward, the spacing 
    between grid points increases from 100 K to 200 K.}
\end{figure*}

This missing-lines phenomenon can arise for three different reasons: Stellar atmospherse genuinely do not produce that line at detectable strength at the given temperature and surface gravity. Second, PHOENIX missed it, possibly accidentally.  Or third, our line-finding and line-association algorithms missed it.  Whatever the cause, these missing lines have immediate practical consequences.  In theory, interpolation schemes may be inaccurate in regions where 
a line does not appear, or could break algorithms that demand a rectilinear grid, such as \texttt{scipy}'s \texttt{RegularGridInterpolator}.

To solve this, we artificially
populated missing grid points with log-amplitudes of -1000, which retained
interpolator stability by not being infinity, but also essentially 
nullified the line when evaluated. An example of the appearance of missing 
sections in heatmaps where a line does not appear is shown in Figure 3. 
In total, across the entire PHOENIX subset, \texttt{blase} detected 
128,723 individual spectral lines.

Every one of these lines can be visualized as a manifold: a mapping between one 4-dimensional heatmap of line properties to a 3-dimensional heatmap of fundamental properties. All of these 128,723 manifolds have to be interpolated.

@@

\subsection{Continuously Evaluable Manifolds}
For each line, the inputs to the interpolator were the three input parameters $T_{\mathrm{eff}}$,
$\log(g)$, and [Fe/H], and the output was a list of four parameters, 
$\mu'$, $\ln(a)$, $\sigma$, and $\gamma$. For each line, one of these 
interpolator objects was created using linear interpolation, and these 
interpolators were aggregated into a single list, which was then
written to disk in the \texttt{.pkl} file storage format.  These interpolators generate multiple 
manifolds with the following mapping:
\begin{gather}
    \mqty[T_{\mathrm{eff}} \\ \log(g) \\ \mathrm{[Fe/H]}] \rightarrow \mqty[\mu' \\ \ln(a) \\ \sigma \\ \gamma]
\end{gather}
These interpolators could now be evaluated at any point lying within the 
domain of the PHOENIX subset, turning a discretely sampled PHOENIX subset 
into a continuously evaluable function, sometimes called a spectral emulator or surrogate model. With the given size of the PHOENIX subset, the 
interpolator list takes up 13.2 GB of disk space. This evaluation 
is able to reconstruct an existing PHOENIX spectrum or alternatively 
interpolate a new spectrum within the domain of the PHOENIX subset,
we call this the PHOENIX generator. In Figure 4, we show the same 
spectral line as in Figure 3, but now supersampled using the PHOENIX
generator evaluated over the same slice.

The spectral reconstruction process is done by iterating over the 
PHOENIX generator, evaluating each interpolator at the given coordinates, 
then reshaping the data into the same format that PyTorch uses for 
state dictionaries. During the iteration, if the interpolated
log-amplitude of the line is less than -100, the line is excluded from the 
state dictionary. The rationale for -100 instead of -1000 originates from the fact that at wavelength regions near artificially populated lines, the manifold experiences artifacts due 
a sudden decline. We have to anticipate that decline and cut off the 
behavior in advance. This saves on computational time. 
\authorcomment1{I do not understand this.  Maybe consider removing?}

Finally, the state dictionary is fed into \texttt{blase}'s \texttt{SparseLinearEmulator},
which reconstructs the spectrum by constructing a forward model based on 
the input state dictionary. Any \texttt{nan} values are set to 1 (which we 
can do because the spectra are all normalized), and the spectrum is 
complete.

@@

\section{Bayesian Inference and Testing}
\subsection{Spectral Reconstruction Time}
The typical use case for the PHOENIX generator may be to batch 
reconstruct spectra from an array of input coordinates \authorcomment1{which input coordinates-- fundamental properties or line properties?}. Therefore, there
is some motivation to reduce the computational cost of this
procedure to tractable levels. We evaluate the computational time needed
to use the PHOENIX generator in two distinct ways. First, for a single
input, which would be relevant in serial applications. Second, for an array
of multiple inputs, which would be relevant in parallel applications.
The \texttt{RegularGridInterpolator} API allows for the passing in of 
an entire array of input coordinates to be evaluated at once, however 
using \texttt{blase} to reconstruct the spectrum from our interpolated
state dictionary is done serially, leading to what is actually more of 
a psuedo-parallel evaluation. Performance results are shown in Figure 5,
and we can see that the pseudo-parallel implementation is much faster.

@@

\subsection{Inference Algorithm}
We elected to use Bayesian optimization as the inference algorithm, specifically 
the \texttt{gp\_minimize} function from the \texttt{scikit-optimize} library \citep{skopt}. 
This algorithm uses a Gaussian Process to model the objective function, which in 
this case was the RMS (Root-Mean-Square) loss between the interpolated spectrum 
$\mathsf{M}$ and the true spectrum $\mathsf{D}$, defined as:
\begin{gather}
    \mathcal{L} = \ev{(\mathsf{M} - \mathsf{D})^{\circ 2}}^{1/2}
\end{gather}
The optimizer was configured to first run 100 random 
evaluations to seed the surrogate model, then run 20 more evaluations now guided 
by the surrogate model. This totals to 120 evaluations, which was deemed 
sufficient for this study. One inference run takes on average just under 
7.5 minutes to complete.

\subsection{Bayesian Optimizer Performance}
To test the performance of the inference algorithm, we used the PHOENIX subset
itself. At first glance, this may seem circular, as the 
PHOENIX generator has memorized the PHOENIX subset, being able to
reconstruct a PHOENIX spectrum when evaluated at that grid point. However,
that stategy allows us to use the PHOENIX subset as test data.
The Bayesian optimizer's surrogate model is seeded by random continuously 
sampled generator evaluations within the search space, \textit{not} grid 
points of the PHOENIX subset, meaning the surrogate model has no `memorization' 
to speak of. If the optimizer had been an iterative grid refinement search, 
this would not have been possible, because grid methods rely on just that, 
the grid.

We know that in typical observational settings, a coarse estimate for $T_{\mathrm{eff}}$ tends to be fairly well-constrained from ancillary data prior to 
even looking at the spectra, so when testing the inference algorithm, we limited
its search space to only include $T_{\mathrm{eff}}$ which lay within 500 K of the
true value on either side. $\log(g)$ and [Fe/H] were allowed to vary freely. 
The test sample consisted of 9 unique $T_{\mathrm{eff}}$ values, 3 unique $\log(g)$ 
values, and 2 unique [Fe/H] values, totaling 54 unique spectra. $T_{\mathrm{eff}}$
ranged from 3000 K to 11000 K in increments of 1000 K, $\log(g)$ ranged from 2 to 6
in increments of 2, and [Fe/H] ranged from -0.5 to 0 in increments of 0.5.

@@

\section{Final Considerations}
The computer used for this study, Triton, has the following specifications:
\begin{table}[h!]
    \centering
    \begin{tabular}{ll}
        \hline
        CPU & AMD EPYC 7513\\
        RAM & 256 GB\\
        GPU & Nvidia A100 40GB ($\times$2)\\
        \hline
    \end{tabular}
    \caption{This machine was used for all computations, but not 
    for generating visualizations. The EPYC 7513 is a 32c/64t CPU with a 
    boost clock of 3.65 GHz. The A100 has 6912 CUDA cores.}
\end{table}

The interpolation scheme presented here represents a proof of concept, showing that leveraging the mapping between synthetic spectral lines and their inputs can yield a semi-empirical basis for data-model comparisons.  There are numerous design considerations that could be improved upon with 

@@

    study did not include the full PHOENIX grid, which expands 

@@

    and also includes the alpha element abundance parameter, which we 
